Applying Spatio-temporal Scan Statistics and Spatial Autocorrelation Statistics to identify Covid-19 clusters in the world.

### 2021-01-13 Notebooks/paper_covid.rmd
Update paper_covid.rmd

 Both the Poisson Scan Statistic proposed by Kulldorff, the Global Moran's I and the *LISA* were already used before in studies about spatial and spatio-temporal characteristics of Sars-CoV-2, the Scan Statistic for instance was used once in studies about the spatio-temporal behavior of Covid-19 in the state of Sergipe in Brazil, while both the Global and Local Moran's I were used in chinese studies regarding the spatial characteristics of the pandemic in China, here these methods showed themselves to be useful while studying the spatial and spatio-temporal characteristics of the virus.

 While the Global Moran's I could be used with great success to identify spatial association in smaller areas affected by the virus and help in decision making while dealing with local outbreaks, the *LISA* would be useful not only to identify with more precision where the spatial association is occurring but it also should be used when the goal is to control Covid-19 outbreaks in larger areas. The Spatio-Temporal Scan Statistic should be used when trying to identify "alive" *clusters* and their initial date, therefore it would be a great tool in decision making when dealing not only with Covid-19 but with other disease outbreaks.

>>>>

 Both the Poisson Scan Statistic proposed by Kulldorff, the Global and the Local Moran's I were already used before in studies about spatial and Spatio-temporal characteristics of Sars-CoV-2, whereas the Scan Statistic for instance was used once in studies about the Spatio-temporal behavior of Covid-19 in the state of Sergipe in Brazil, while both the Global Moran's I and the LISA were used in Chinese studies regarding the spatial characteristics of the pandemic in China, in this paper these methods showed themselves to be useful when there is a need to study spatial and Spatio-temporal characteristics of a virus.

 While the Global Moran's I could be used with great success to identify the spatial association in smaller areas affected by the virus and help in decision making while dealing with local outbreaks, the LISA would be useful not only to identify with more precision where the spatial association is occurring but it also should be used when the goal is to control Covid-19 outbreaks in larger areas. The Spatio-Temporal Scan Statistic should be used when trying to identify "alive" clusters and their initial date, therefore it would be a great tool in decision making when dealing not only with Covid-19 but with other disease outbreaks.
